\vspace{-2.25cm} Multimodal Abnormity-Aware Model\\for Ocular Disease Diagnosis

### 2024-08-28 Paper/Paper.tex
Get illustration PPT out; draw OCT and Fundus source tables

	We develop the Multimodal Abnormity-Aware Model (MAAM) to aid ophthalmologists in diagnosing ocular diseases. It considers both OCT and color fundus image inputs. In addition, it simulates the decision-making process of ophthalmologists when they examine patients, i.e., first identifying the abnormities, and then deducing the disease. The model is designed in a way that partly opens up the ``black box'' of DL and presents more interpretable information for the reference of the ophthalmologists.

@@

	The structure of the MAAM is shown in Fig.~\ref{fig:3_parts}. It consists of 2 submodels: the Abnormity Models and the Diagnosis Model. We input multiple OCT and fundus images to Abnormity Models, and output the classification of abnormities as inputs to Diagnosis Model, which generates disease probabilities as final result.

@@

	The Abnormity Models also contains 2 submodels: the OCT Model and the Fundus Model, which identify OCT and fundus abnormities, respectively. Both of the submodels leverage CNNs with modified final full-connected (FC) layers. We compare the performances of 4 commonly used CNN, ResNet152, ResNet50, ResNet18 and VGG16, and choose the best one for each submodel. The final FC layers in each CNN are modified so that the size of their output vectors is equal to the number of OCT or fundus abnormities. In order to do normalization, we perform softmax operation for the FC layer. After an image is inputted into one of the submodels, the submodel returns a probability vector for all the abnormities.

@@

	And the Diagnosis Model consists of two stages: the Stage D1 and Stage D2.

@@

	In Stage D1, we determine the severity level for each disease from the probability vectors from the Abnormity Models.  We use Abnormity to Disease Deduction Criteria (shown in Fig.~\ref{fig:criteria}) as ground truth. There are multiple submodels in Stage D1 and each one corresponds to one disease. From the deduction criteria, we determine the number of abnormities for one disease and use the number to define the severity levels. For example, there are totally 5 abnormities that are present in disease ``Referable DR'', i.e., OCT abnormity ``DME'' and Fundus abnormities ``Hemorrhage'', ``Vascular Abnormity'', ``Microaneurysm'', and ``Cotton Wool Patch''.  Therefore, when taking account of healthy status, we end up with 6 severity levels for disease ``Referable DR''.

	We use fused vector as an input for each submodels in Stage D1, and have the vector pass a FC layer with softmax operation to generate severity level probability vectors.

@@

	In Stage D2, we determine the final disease probability vector.  Similarly we use a fused vector from the output of submodels in Stage D1, and have the vector pass a FC layer with softmax operation.  The result presents potential diseases that can be referenced by ophthalmologist.

@@

	In MAAM, we leverage the fusion operation to incorporate all the results from different submodels.  As shown in Fig.~\ref{fig:fusion}, there are 3 fusion operations in MAAM.

@@

	The first fusion occurs at the output of OCT Models or Fundus Models.  In reality, multiple OCT images and Fundus images can be used for ocular disease diagnosis. Each image goes through one of the Abnormity Models and generates a probability vector. To use the probability results from all the images, a maximization fusion is implemented for all the vectors so that the fusion yeilds one OCT abnormity probability vector and one Fundus abnormity probability vector with maximum values.

	The second fusion occurs at the interface between Abnormity Model and Stage D1. A concatenation fusion is implemented for OCT abnormality probability vector and Fundus abnormality probability vector.  The fused vector works as an input to all the submodels in Stage D1.

	The third fusion occurs at the interface between Stage D1 and Stage D2.  We concatenation all the severity level vectors to feed the model in Stage D2.  The fused vector goes through the FC layer and does a softmax operation to yield the final result.

@@

	Table: Data source for OCT and Fundus
	numbered list of sources
	columns for source number, final column = total number of imgs
	rows for abnormities

	TEXT: CycleGAN
	Table: selection rates
	selection criteria
	Why I did not GAN Fundus

@@

	Table: all data source

@@

	{
		\fontsize{9}{12}\selectfont
		{
			\begin{longtable}{lp{3.8in}}
				\caption{OCT Abnormities}
				\label{tb:oct-abnormites}\\
				\toprule
				Abnormality&Description\\
				\toprule

@@

	{
		\fontsize{9}{12}\selectfont
		{
			\begin{longtable}{lp{3.8in}}
				\caption{Fundus Abnormalities}
				\label{tb:fundus-ab}\\
				\toprule
				Abnormality&Description\\
				\toprule

@@

				\multicolumn{1}{l}{Vascular abnormality} & Retinal Vessel Tortuosity.  Retinal Vessel Caliber Changes.  \\

>>>>

	We develop the Multimodal Abnormity-Aware Model (MAAM) to aid ophthalmologists in diagnosing ocular diseases. It considers both OCT and color fundus image inputs. In addition, it simulates the decision-making process of ophthalmologists when they examine patients, i.e., first identifying the abnormities, and then deducing the disease. The model is designed in a way that partly opens up the ``black box'' of DL and presents more interpretable information for the reference of ophthalmologists.

@@

	The structure of the MAAM is shown in Fig.~\ref{fig:3_parts}. It consists of 2 submodels: the Abnormity Models and the Diagnosis Model. We input multiple OCT and fundus images to the Abnormity Models, and output the classification of abnormities as inputs to the Diagnosis Model, which generates disease probabilities as the final result.

@@

	The Abnormity Models also contain 2 submodels: the OCT Model and the Fundus Model, which identify OCT and fundus abnormities, respectively. Both of the submodels leverage CNNs with modified final fully-connected (FC) layers. We compare the performances of 4 commonly used CNNs, ResNet152, ResNet50, ResNet18 and VGG16, and choose the best one for each submodel. The final FC layer in each CNN is modified so that the size of its output vector is equal to the number of OCT or fundus abnormities. In order to normalize the output vector, we perform the softmax operation. After an image is inputted into one of the submodels, the submodel returns a probability vector for all the abnormities.

@@

	The Diagnosis Model consists of two stages: Stage D1 and Stage D2.

@@

	In Stage D1, we determine the severity level for each disease from the probability vectors from the Abnormity Models. We use Abnormity-to-Disease Deduction Criteria (shown in Fig.~\ref{fig:criteria}) as ground truth. There are multiple submodels in Stage D1 and each one corresponds to one disease. From the deduction criteria, we determine the number of abnormities for one disease and use the number to define the severity levels. For example, there are totally 5 abnormities that are present in the disease ``Referable DR'': OCT abnormity ``DME'' and Fundus abnormities ``Hemorrhage'', ``Vascular Abnormity'', ``Microaneurysm'', and ``Cotton Wool Patch''. Therefore, taking into account the healthy status, we end up with 6 severity levels for disease ``Referable DR''.

	We use the fused vector as an input for each submodels in Stage D1, and have the vector pass a FC layer with softmax operation to generate severity level probability vectors.

@@

	In Stage D2, we determine the final disease probability vector. Similarly we use a fused vector from the outputs of submodels in Stage D1, and have the vector pass a FC layer with softmax operation. The result presents potential diseases that can be referenced by ophthalmologists.

@@

	In the MAAM, we leverage the fusion operation to incorporate all the results from different submodels. As shown in Fig.~\ref{fig:fusion}, there are 3 fusion operations in MAAM.

@@

	The first fusion occurs at the output of OCT Model or Fundus Model. In real cases, multiple OCT and Fundus images can be used for ocular disease diagnosis. Each image goes through one of the Abnormity Models and generates a probability vector. To use the probability results from all the images, a maximization fusion is implemented for all the vectors so that the fusion yields one OCT abnormity probability vector and one Fundus abnormity probability vector with maximum values.

	The second fusion occurs at the interface between Abnormity Model and Stage D1. A concatenation fusion is implemented for OCT abnormity probability vector and Fundus abnormity probability vector. The fused vector works as an input to all the submodels in Stage D1.

	The third fusion occurs at the interface between Stage D1 and Stage D2. We concatenation all the severity level vectors to feed the model in Stage D2. The fused vector goes through the FC layer and undergoes a softmax operation to yield the final result.

@@

	The data preparation process is shown in Fig.~\ref{fig:A_data_prep}. First,

	\begin{minipage}[t]{0.4\linewidth}
		{
			\fontsize{9}{12}\selectfont
			{
				\begin{longtable}{cccccc}
					\caption{OCT abnormities}
					\label{tb:OCT_source}\\
					\toprule
					\multirow{2}{*}{Abnormity}&\multicolumn{4}{c}{Source}&\multirow{2}{*}{Total}\\
					&1&2&3&4&\\
					\midrule
					CNV               &2984&/  &/ &/ &2984\\
					CSC               &/   &102&32&/ &134 \\
					DME               &2500&/  &/ &/ &2500\\
					Drusen            &2500&/  &/ &/ &2500\\
					ERM               &/   &/  &/ &19&19  \\
					Intraretinal&1097&/  &/ &/ &1097\\
					MH                &/   &99 &31&/ &130 \\
					Mactel            &/   &/  &29&/ &29  \\
					Normal            &5000&/  &/ &/ &5000\\
					RP                &/   &102&31&/ &133 \\
					Stargardt         &/   &/  &23&/ &23  \\
					Subretinal  &1083&/  &/ &/ &1083\\
					\bottomrule
				\end{longtable}

				\vspace{0.5cm}
				\begin{enumerate}
					\item Normal Disease Database
					\vspace{-0.2cm}

					\item OpenICPSR
					\vspace{-0.2cm}

					\item Few-shot
					\vspace{-0.2cm}

					\item Google Image
					\vspace{-0.2cm}
				\end{enumerate}

				\vspace{0.5cm}
			}
		}
	\end{minipage}
	\begin{minipage}[t]{0.6\linewidth}
		{
			\fontsize{9}{12}\selectfont
			{
				\begin{longtable}{cccccccc}
					\caption{Fundus abnormities}
					\label{tb:Fundus_source}\\
					\toprule
					\multirow{2}{*}{Abnormity}&\multicolumn{6}{c}{Source}&\multirow{2}{*}{Total}\\
					&1&2&3&4&5&6&\\
					\midrule
					Cotton  &/  &/ &/ &33 &205&/ &238\\
					Drusen            &/  &/ &/ &50 &/  &/ &50 \\     
					Hard    &20 &/ &/ &75 &284&/ &379\\ 
					Hemorrhage        &13 &66&/ &105&278&/ &462\\     
					MH                &/  &/ &/ &/  &/  &34&34 \\        
					Micro   &55 &/ &/ &1  &219&/ &275\\
					Normal            &100&37&15&/  &/  &/ &152\\      
					RP                &/  &22&/ &/  &/  &44&66 \\         
					Vascular&/  &64&/ &14 &/  &/ &78 \\

					\bottomrule
				\end{longtable}

				\vspace{1cm}
				\begin{enumerate}

					\item E-ophtha
					\vspace{-0.2cm}

					\item Kaggle1000
					\vspace{-0.2cm}

					\item HRF
					\vspace{-0.2cm}

					\item STARE
					\vspace{-0.2cm}

					\item EyePACS
					\vspace{-0.2cm}

					\item Google Image
					\vspace{-0.2cm}

				\end{enumerate}

				\vspace{0.5cm}
			}
		}
	\end{minipage}

	{
		\fontsize{9}{12}\selectfont
		{
			\begin{longtable}{ccccc}
				\caption{OCT images CycleGAN selection rate}
				\label{tb:cycleGAN_selection_rate}\\
				\toprule
				Abnormity&Original&Generated&Selected&Selection Rate\\
				\midrule
				CSC      &134&1730&3000&57.667\% \\
				ERM      &19 &1931&3000&64.367\% \\
				MH       &130&1811&3000&60.367\% \\
				Mactel   &29 &1900&3000&63.333\% \\
				RP       &133&1923&3000&64.100\% \\
				Stargardt&23 &2582&3000&86.067\% \\
				\bottomrule
			\end{longtable}
		}
	}

	TEXT: CycleGAN
	selection criteria
	Why I did not GAN Fundus

@@

@@

	{
		\fontsize{9}{12}\selectfont
		{
			\begin{longtable}{lp{3.8in}}
				\caption{OCT Abnormities}
				\label{tb:oct-abnormites}\\
				\toprule
				Abnormity&Description\\
				\toprule

@@

	{
		\fontsize{9}{12}\selectfont
		{
			\begin{longtable}{lp{3.8in}}
				\caption{Fundus Abnormalities}
				\label{tb:fundus-ab}\\
				\toprule
				Abnormity&Description\\
				\toprule

@@

				\multicolumn{1}{l}{Vascular abnormity} & Retinal Vessel Tortuosity. Retinal Vessel Caliber Changes. \\
